Coasting on Duality: What We Get When We Do Two-Mode CA

### 2024-07-12 slides/duality-presentation.qmd
changes

---
title: "Coasting on Duality"
subtitle: "What We Get When We Do Two-Mode Network CA"
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
slide-number: true
slide-level: 2
progress: true
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    footer: "https://github.com/olizardo/correspondence-analysis-two-mode-networks"
---

>>>>

---
title: "Coasting on Duality"
subtitle: "What We Get When We Do Two-Mode Network CA"
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
format:
  beamer:
    incremental: false 
    theme: simple 
    smaller: true
    slide-number: true
    slide-level: 2
    progress: true
    footer: "https://github.com/olizardo/correspondence-analysis-two-mode-networks"
---

### 2024-07-12 slides/ncsu-presentation.qmd
changes

---
title: "Coasting on Duality: What We Get When We Do Two-Mode CA"
author: "Omar Lizardo (UCLA)"
format:
  revealjs:
    incremental: false 
    theme: sky 
    smaller: true
---

- CA Proponents point to some important advantages of CA for the analysis of two-mode networks:
  - CA can provide synoptic graphical representations of structural connectivity patterns in two-mode data**
  - (M)CA can be used to "visually explore" two-mode networks***

**Breiger 2000; Roberts 2000; Faust 2005

***D’Esposito et al. 2014
:::

- This presentation builds on and expands on previous  research on using CA for two-mode data analysis

1. CA as computing a type of "(Reflective) Dual Centrality" on the entities in the two-modes**
    - Rediscovered recently under the heading of the Economic Complexity Index (ECI; Hidalgo and Hausman 2009)

@@

2. CA as extracting "positional" information the entities in the two modes
    - With positions defined by "Generalized Relational Similarity"**
    - Similar entities have similar patterns of connectivity with entities that are themselves similar

@@

*Kovacs 2010; Lizardo 2024
:::

@@

::: aside
*Later referred to as the Economic Complexity Index (ECI)

@@

## Southern Women Data {.scrollable}
```{r Southern women data}
    A <- matrix(
         c(
          1, 1, 1, 1, 1,  1, 0, 1, 1,  0, 0, 0, 0, 0, 
          1, 1, 1, 0, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 1, 1, 1, 1,  1, 1, 1, 1,  0, 0, 0, 0, 0, 
          1, 0, 1, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  0, 1, 0, 0,  0, 0, 0, 0, 0, 
          0, 0, 1, 1, 1,  1, 0, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  1, 1, 1, 0,  0, 0, 0, 0, 0, 
          0, 0, 0, 1, 1,  0, 1, 1, 1,  0, 0, 0, 0, 0,

          0, 0, 0, 0, 0,  0, 1, 1, 1,  0, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 0, 0, 
          0, 0, 0, 0, 0,  0, 0, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 1,  1, 0, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 0, 1,  1, 1, 1, 1, 1,
          0, 0, 0, 0, 0,  0, 1, 1, 0,  1, 1, 1, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 0, 1,  0, 1, 0, 0, 0,

          0, 0, 0, 0, 0,  1, 0, 1, 1,  0, 0, 0, 0, 0,
          0, 0, 0, 0, 0,  0, 0, 1, 1,  0, 0, 0, 0, 0),
          ncol = 14, byrow = TRUE)
     w <- c("EVELYN", "LAURA", "THERESA", "BRENDA", "CHARLOTTE", "FRANCES", "ELEANOR", "RUTH", "VERNE", "MYRA", "KATHERINE", "SYLVIA", "NORA", "HELEN", "OLIVIA", "FLORA", "PEARL", "DOROTHY")
     e <- paste("E", c(1:14), sep = "")
     rownames(A) <- w
     colnames(A) <- e
     require(kableExtra)
     A <- A[sort(rownames(A)), ]
     kbl(A)
```

@@

## CA and Bonacich Centrality
- But what's the connection to reflections and CA after all of this?

- A couple of things first:

  - $Dp$ is a $|P| \times |P|$ matrix containing each person's $C_p^R(1)$ (activity) along the diagonals and zeroes in every other cell
  - $Dg$ is a $|G| \times |G|$ matrix $Dg$ containing each group's $C_g^R(1)$ (size) along the diagonals and zeroes in every other cell

- In the limit ($q \rightarrow \infty$), the iterative HH reflective centralities are the solution of the eigensystem:*

$$
    \left(Dp^{-1}ADg^{-1}A^T\right)C^R_p = \lambda C^R_p
$$ {#eq-dam1}

$$
    \left(Dg^{-1}A^TDp^{-1}A\right)C^R_g = \lambda C^R_g
$$ {#eq-dam2}

::: aside
*van Dam et al. 2021
:::

@@

@@

::: {.notes}
As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups. 
:::

@@

@@

- In *SimRank* the similarity between two people is given by:

$$
S(p, p') = \frac{\alpha}{C_p^R(1)C_{p'}^R(1)}
    \sum_{i = 1}^{C_p^R(1)} \sum_{j = 1}^{C_{p'}^R(1)} 
    S\left(g(i)_{i \in N(p)}, g(j)_{j \in N(p')}\right)
$$ {#eq-simrank1}

>>>>

---
title: "Coasting on Duality: What We Get When We Do Two-Mode CA"
author: "Omar Lizardo (UCLA)"
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
---

@@

@@

@@

::: aside
*Referred to as the Economic Complexity Index (ECI)

@@

@@

@@

::: aside
*Mealy 2019, eq. 4
:::

@@

::: {.notes}
As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups. 
:::
 be thought of as **similarity matrices** based on the shared neighbors in the two mode network
## CA Reveals Dual Communities
- As we know from Breiger (1974) the usual one mode projections are obtained via:

@@

- Which can

@@
